Testing for measurement invariance with respect to an ordinal variable

### 2012-08-21 ordinal/ordinal.Rnw
improved some comments
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/semtools@163 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

%% TODO add some more citations to these initial paragraphs
The study of measurement invariance and differential item functioning
(DIF) 
has received considerable attention in the psychometric literature
(see, e.g., \citeNP{Mil11} for a thorough review).  A set of
psychometric scales ${\bm X}$ are defined to be measurement invariant
with respect to an auxiliary variable $V$ if \cite{Mel89}
\begin{equation}
    \label{eq:midef}
      f(x_i | t_i, v_i, \dots) = f({\bf x}_i | t_i, \dots),
\end{equation}
where $T$ is the latent variable that the scales measure, $f$ is the
model's distributional form, and the $i$ subscript refers to
individual cases.  If the above equation does not hold, then there
exists a measurement invariance violation.  We focus here on
situations where $f()$ is parametric, and the measurement invariance
violation occurs because the model parameters are
unequal across individuals (and related to $V$).
\readme{Maybe we should explicitly mention that in parametric settings, f is
the probability density function of x? Also, do we need to point out
the relation between X and x, V and v, T and t?}

@@

\readme{My feeling is that we still need to explain better what type
of patterns are covered by which alternative. Maybe we should do it
the other way round and start with $H_2^*$ as the most general type
of deviation from $H_0$. Then, we could explain that these also
covers changes that are not `monotonic' in V, i.e., where parameter
values change in one direction as V increases and then change into
another direction again. After that we could continue and argue that
such monotonic changes are of particular interest for categorical V
and that the simplest form of such a change is $H_1^*$. And as a final
note one could add that this one-shift alternative is also often
employed in practice where practitioners often use the median V
as the cutoff whereas we want to assess all possible cutoffs.}

>>>>

%% TODO add some more citations to these initial paragraphs
The study of measurement invariance and differential item functioning
(DIF) 
has received considerable attention in the psychometric literature
(see, e.g., \citeNP{Mil11} for a thorough review).  A set of
psychometric scales ${\bm X}$ are defined to be measurement invariant
with respect to an auxiliary variable $V$ if \cite{Mel89}
\begin{equation}
    \label{eq:midef}
      f(x_i | t_i, v_i, \dots) = f({\bf x}_i | t_i, \dots),
\end{equation}
where $T$ is the latent variable that the scales measure, $f$ is the
model's distributional form, and the $i$ subscript refers to
individual cases.  If the above equation does not hold, then there
exists a measurement invariance violation.  We focus here on
situations where $f()$ is parametric, and the measurement invariance
violation occurs because the model parameters are
unequal across individuals (and related to $V$).
\fixme{Maybe we should explicitly mention that in parametric settings, f is
the probability density function of x? Also, do we need to point out
the relation between X and x, V and v, T and t?}

@@

\fixme{My feeling is that we still need to explain better what type
of patterns are covered by which alternative. Maybe we should do it
the other way round and start with $H_2^*$ as the most general type
of deviation from $H_0$. Then, we could explain that these also
covers changes that are not `monotonic' in V, i.e., where parameter
values change in one direction as V increases and then change into
another direction again. After that we could continue and argue that
such monotonic changes are of particular interest for categorical V
and that the simplest form of such a change is $H_1^*$. And as a final
note one could add that this one-shift alternative is also often
employed in practice where practitioners often use the median V
as the cutoff whereas we want to assess all possible cutoffs.}
